---
title: "Lab1_Practice"
author: "Natalie Smith"
format:
  html:
    code-fold: true
    toc: true
excecute: 
  echo: true
  message: false
  warning: false
editor: visual
---

```{r}
library(tidyverse)
library(gapminder)
```

Try on your own:

In your github account, create a new repository called "lab1_practice", then clone it to your laptop.

In the new R project:

1.  Create a new quarto document. If you like, set the header for code folding. Set echo message, and warning like we did in lab. (X)
2.  Load tidyverse and gapminder. (X)
3.  For each year and continent, summarize the average per-capita GDP across all countries in the continent. Hint: don't just average the country level values - need to consider country populations!

```{r}
data("gapminder")
gapminder_av_gdp <-gapminder %>% 
  group_by(continent, year) %>% summarize(average_per_capita_gdp = mean(gdpPercap))

gapminder_av_gdp
```

4.  Plot the continent-level GDP values in a scatter plot. Bonus, plot the country-levels as well, with a low alpha to fade them into the background.

```{r}
gdp_scatterplot <- ggplot(data=gapminder_av_gdp, aes(x= year, y= average_per_capita_gdp, color= continent))+geom_point()+
  labs(y = "Average Per Capita GDP", x="Year", color = "Continent")+
  ggtitle("Continent-Level GDP Value by Year ")+
  theme_minimal()

gdp_scatterplot

```

5.  Redo the scatter plot bu this time, log transform the GDP data (either in the data frame using 'mutate', or on the plot with 'scale_y_log10()')

```{r}
gdp_scatterplot_log <- ggplot(data=gapminder_av_gdp, aes(x= year, y= average_per_capita_gdp, color= continent))+geom_point()+
  scale_y_log10()+
  labs(y = "Average Per Capita GDP (log10)", x="Year", color = "Continent")+
  ggtitle("Continent-Level GDP Value by Year ")+
  geom_smooth(method="lm",se=FALSE)+
  theme_minimal()
  
gdp_scatterplot_log 
```

Based on the plots, choose either regular or log transformed data, and use a linear regression to estimate annual change in per capita GDP \*\* OR \*\* log(per capita GDP) accounting for continent.

```{r}
lm_gdp_scatterplot_log <- ggplot(data=gapminder_av_gdp, aes(x= year, y= average_per_capita_gdp, color= continent))+geom_point()+
  scale_y_log10()+
  labs(y = "Average Per Capita GDP (log10)", x="Year", color = "Continent")+
  ggtitle("Continent-Level GDP Value by Year ")+
  theme_minimal()+
  geom_smooth(method="lm",se=FALSE)
  
lm_gdp_scatterplot_log

```

```{r}
linear_model <- lm((average_per_capita_gdp) ~ continent, data = gapminder_av_gdp)

summary(linear_model)
```

-   Write a sentence or two describing the trends in plots. Average per capita GDP is increasing across all continents from 1952 - 2007. Oceania maintains the highest GDP.

-   Write a sentence or two to explain to coefficients in the linear model, including statistical significance. Measuring the change in average GDP as compared to Africa. They all have a p-value lower than 0.05 meaning they are significantly different than Africa.

For every change in x by 1, you get 16428 change in Oceania.

-   Which continent is the reference continent? Why might that be? Africa - alphabetical order

-   What do the coefficients on each continent represent? Do these make intuitive sense? Change in slope - so for every 1 change in x you get \_\_\_\_change in y based on linear output.

-   Does a linear model seem like a good way to approach these data? Why or why not? Maybe - depending on the question you are asking. Test assumptions (normality, heteroscedasticy, independent samples)

When done render your Quatro doc, then stage, commit, and push back to Github.

Congrats, you did it!
